---
title: "consensus HLA Report"
output:
  pdf_document:
    highlight: tango
number_sections: TRUE
params:
  clin_hla_json: NAMESPACE
  germline_hla_json: NAMESPACE
  tumour_hla_json: NAMESPACE
  rna_hla_json: NULL
  pid: NAMESPACE
date: "`r format(Sys.time(), '%d %B, %Y')`"
subtitle: "`r paste('Patient ', params$pid)`"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(rjson)
library(dplyr)
library(knitr)
library(kableExtra)
library(stringr)

date <- format(Sys.time(), "%d %B, %Y")
```


This HLA report was generated using the consHLA toolkit on `r date`.

# Summary of Clinically Relevant HLA Consensus Alleles
```{r echo = FALSE}

hla_from_json <- rjson::fromJSON(file = params$clin_hla_json)
hla_df <- as.data.frame(hla_from_json)

small_hla_df <- hla_df %>%
  select(
    HLA.A, HLA.B, HLA.C, HLA.DRA, HLA.DRB1, HLA.DRB3, HLA.DRB4,
    HLA.DRB5, HLA.DQA1, HLA.DQB1, HLA.DPA1, HLA.DPB1
  )

cols <- c("HLA-A", "HLA-B", "HLA-C", "HLA-DRA", "HLA-DRB1", "HLA-DRB3", "HLA-DRB4", "HLA-DRB5", "HLA-DQA1", "HLA-DQB1", "HLA-DPA1", "HLA-DPB1") # nolint
colnames(small_hla_df) <- c("HLA-A", "HLA-B", "HLA-C", "HLA-DRA", "HLA-DRB1", "HLA-DRB3", "HLA-DRB4", "HLA-DRB5", "HLA-DQA1", "HLA-DQB1", "HLA-DPA1", "HLA-DPB1") # nolint

class_i <- small_hla_df[, c("HLA-A", "HLA-B", "HLA-C")]
class_ii_a <- small_hla_df[, c("HLA-DRA", "HLA-DRB1", "HLA-DRB3", "HLA-DRB4", "HLA-DRB5")] # nolint
class_ii_b <- small_hla_df[, c("HLA-DQA1", "HLA-DQB1", "HLA-DPA1", "HLA-DPB1")] # nolint

```

## Class-I
```{r}
kable(class_i, "latex", booktabs = TRUE, escape = FALSE, digits = 2, col.names = linebreak(c("HLA-A", "HLA-B", "HLA-C"), align = "c")) %>% # nolint
  kable_styling(latex_options = c("HOLD_position", "striped"), full_width = FALSE, position = "left", font_size = 9) # nolint
```
## Class-II
```{r}
kable(class_ii_a, "latex", booktabs = TRUE, escape = FALSE, digits = 2, col.names = linebreak(c("HLA-DRA", "HLA-DRB1", "HLA-DRB3", "HLA-DRB4", "HLA-DRB5"), align = "c")) %>% # nolint
  kable_styling(latex_options = c("HOLD_position", "striped"), full_width = FALSE,position = "left", font_size = 9) # nolint
kable(class_ii_b, "latex", booktabs = TRUE, escape = FALSE, digits = 2, col.names = linebreak(c("HLA-DQA1", "HLA-DQB1", "HLA-DPA1", "HLA-DPB1"), align = "c")) %>% # nolint
  kable_styling(latex_options = c("HOLD_position", "striped"), full_width = FALSE,position = "left", font_size = 9) # nolint
```

\newpage

# All Alleles
```{r echo = FALSE}
germline_hla_json <- rjson::fromJSON(file = params$germline_hla_json)
tumour_hla_json <- rjson::fromJSON(file = params$tumour_hla_json)
rna_hla_json <- NULL
if (!is.null(params$rna_hla_json)) {
  rna_hla_json <- rjson::fromJSON(file = params$rna_hla_json)
}
truncate_list <- function(x) {
    if (length(x) > 2) {
        return(paste(x[1:2], collapse = ", "))
    } else {
        return(paste(x, collapse = ", "))
    }
}

germline_hla_data <- lapply(germline_hla_json, truncate_list)
tumour_hla_data <- lapply(tumour_hla_json, truncate_list)
rna_hla_data <- if (!is.null(rna_hla_json)) lapply(rna_hla_json, truncate_list) else NULL


germline_hla_df <- as.data.frame(germline_hla_data)
tumour_hla_df <- as.data.frame(tumour_hla_data)
rna_hla_df <- if (!is.null(rna_hla_data)) as.data.frame(rna_hla_data) else NULL


cols <- c("HLA-A", "HLA-B", "HLA-C", "HLA-DRB1", "HLA-DQA1", "HLA-DQB1", "HLA-DPA1", "HLA-DPB1", "HLA-DMA", "HLA-DMB", "HLA-DOA","HLA-DOB","HLA-DRA", "HLA-DRB2", "HLA-DRB3", "HLA-DRB4", "HLA-DRB5", "HLA-DRB6", "HLA-DRB7", "HLA-DRB8", "HLA-DRB9", "HLA-E", "HLA-F", "HLA-G", "HLA-H", "HLA-J", "HLA-K", "HLA-L", "HLA-V") # nolint

colnames(germline_hla_df) <- cols
colnames(tumour_hla_df) <- cols
if (!is.null(rna_hla_df)) colnames(rna_hla_df) <- cols

t_germline_hla_df <- t(germline_hla_df)
t_tumour_hla_df <- t(tumour_hla_df)
t_rna_hla_df <- if (!is.null(rna_hla_df)) t(rna_hla_df) else NULL

swap <- function(df, x, y) {
  temp <- df[x, ]
  df[x, ] <- df[y, ]
  df[y, ] <- temp
  return(df)
}

if (!is.null(t_rna_hla_df)) {
  t_hla_df <- cbind("HLA Gene" = rownames(t_germline_hla_df), t_germline_hla_df, t_tumour_hla_df, t_rna_hla_df)
  
  rownames(t_hla_df) <- 1:nrow(t_hla_df)
  colnames(t_hla_df) <- c("HLA Gene", "Germline WGS", "Tumour WGS", "Tumour RNA-seq")
} else {
  t_hla_df <- cbind("HLA Gene" = rownames(t_germline_hla_df), t_germline_hla_df, t_tumour_hla_df)
  
  rownames(t_hla_df) <- 1:nrow(t_hla_df)
  colnames(t_hla_df) <- c("HLA Gene", "Germline WGS", "Tumour WGS")
}
t_hla_df <- swap(t_hla_df, 9, 13)
t_hla_df <- swap(t_hla_df, 10, 15)
t_hla_df <- swap(t_hla_df, 11, 16)
t_hla_df <- swap(t_hla_df, 12, 17)

determine_consensus <- function(germline, tumour, rna=NULL) {# nolint
  germline_alleles <- sort(unlist(strsplit(germline, ", ")))
  tumour_alleles <- sort(unlist(strsplit(tumour, ", ")))
  if (is.null(rna)) {
    if (grepl("Not typed", germline) || grepl("Not typed", tumour)) {
      return("Not determined")
    }
    if (identical(germline_alleles, tumour_alleles)) {
      return("Full Consensus")
    } else {
      return("Tumour WGS Discrepancy")
    }
  } else {
    rna_alleles <- sort(unlist(strsplit(rna, ", ")))

    if (grepl("Not typed", germline) || grepl("Not typed", tumour) || grepl("Not typed", rna)) {
      return("Not determined")
    }

    if (identical(germline_alleles, tumour_alleles) && identical(tumour_alleles, rna_alleles)) {# nolint
      return("Full Consensus")
    } else if (identical(germline_alleles, tumour_alleles) && !identical(tumour_alleles, rna_alleles)) {# nolint
      return("Tumour RNA-seq Discrepancy")
    } else if (!identical(germline_alleles, tumour_alleles) && identical(tumour_alleles, rna_alleles)) {# nolint
      return("Tumour Consensus")
    } else if (!identical(germline_alleles, tumour_alleles) && !identical(tumour_alleles, rna_alleles) && identical(germline_alleles, rna_alleles)) {# nolint
      return("Tumour WGS Discrepancy")
    } else if (!identical(germline_alleles, tumour_alleles) && !identical(tumour_alleles, rna_alleles) && !identical(germline_alleles, rna_alleles)) {# nolint
      return("Complete Discrepancy")
    } else {
      return("Undefined Discrepancy")
    }
  }
}



add_consensus_column <- function(df) {
  df <- as.data.frame(df)
  
  # Modify to handle optional RNA column
  if (ncol(df) == 4) {
    df <- df %>%
      mutate(consensus = mapply(determine_consensus, df$"Germline WGS", df$"Tumour WGS", df$"Tumour RNA-seq"))
    colnames(df) <- c("HLA Gene", "Germline WGS", "Tumour WGS", "Tumour RNA-seq", "Consensus Status")
  } else {
    df <- df %>%
      mutate(consensus = mapply(determine_consensus, df$"Germline WGS", df$"Tumour WGS"))
    colnames(df) <- c("HLA Gene", "Germline WGS", "Tumour WGS", "Consensus Status")
  }
  
  return(df)
}

df_with_consensus <- add_consensus_column(t_hla_df)
```

```{r}
striped_rows <- seq(1, nrow(df_with_consensus), 2)
kable(df_with_consensus, "latex", booktabs = TRUE, escape = FALSE, digits = 2) %>%
  kable_styling(latex_options = c("HOLD_position", "striped"), full_width = FALSE, position = "left", font_size = 8) %>% # nolint
  row_spec(row = 1:12, bold = TRUE)
```

\newpage

# Appendix
## HLA Typing Method
The consensus HLA typing module uses HLA-HD (v1.4.0) to conduct HLA typing from Tumour and Germline WGS and Tumour RNAseq (optional) for a given patient, producing two or three sets of candidate alleles. For each typed HLA gene, a pair of consensus alleles is calculated based on concordance between the candidate allele sets. 

The first page of the HLA report displays a "clinically significant" subset of HLA alleles in a format which mimics that of a clinical HLA report and is therefore familiar to clinicians. 
The "clinically significant" HLA subset includes most of the protein coding Class I and II HLA genes which are implicated in peptide presentation to T cells while excluding HLA Class I pseudogenes (HLA-H, -J, -K, -L, -V), non-classical Class I genes (HLA-E, -F, -G), Class II pseudogenes (HLA-DRB2, -DRB6, -DRB7, -DRB8, and -DRB9), and the non-classical Class II genes (HLA-DOA, -DOB, -DMA and -DMB). The second page of the report displays a table of all HLA alleles typed by the consHLA module (using HLA-HD), including the clinically relevant subset which is highlighted in bold text.  